---
title: "Data Analysis Project"
author: "Adam Lipman, Logan Beebe, Riley Bomgardner"
date: "5/14/26"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Data Link: <https://www.kaggle.com/datasets/andrewmvd/board-games?resource=download>

## Load libraries

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
library(ggplot2)
library(haven)
```

## Loading Data

```{r}
board_games_data <- read.csv2("data/bgg_dataset.csv") |> clean_names()
```

```{r}
board_games_data
```

## Basic Analysis

There are 20,343 rows and 14 columns in the dataset.

We do not have any price data included in this dataset so we will need joins to other datasets to address that. We also need information about attention deficits and ages that play the most. The column for game mechanics in our data, called mechanics, has inconsistent formatting for separate features, using slashes, commas or spaces to divide them all. They are in the same row which makes it more challenging.

Our dataset does not have any timeframe limitations but some older games like marbles have rough approximates that use negative values. Those values are not specified to be time since the present or if it is in a BCE format. The last year with data is 2022, removing any conclusions from the previous three years. The play_time is not the actual length needed to play the game but instead it is biased data from the game's creators. The reviews for the games do not have sources, making it unclear if they are from one source or combined across multiple, which could have different rating criteria.

We plan to research the needed data categories and try to add that for our future deliverables.

```{r}
nrow(board_games_data)
```

Question: How many versions of Monopoly are there and what affects the different pricing for each?

```{r}
only_monopoly <- board_games_data |> filter(str_detect(name, "Monopoly"))
only_monopoly
```

```{r}
nrow(only_monopoly)
```

```{r}
isolated_mechanics <- only_monopoly |> separate_longer_delim(cols = "mechanics", delim=",")
```

```{r}
counted_mechanics <- isolated_mechanics |> group_by(mechanics) |> summarise(count=n()) |> filter(count==1)
```

```{r}
left_join(isolated_mechanics, counted_mechanics) |> filter(count==1)
```

```{r}
amazon_prices <- read_csv("data/boardgame-geek-dataset_organized.csv")
amazon_prices
```

Question: Does board game popularity correlate with the number of exclusive game mechanics within it?

Findings: There is a slight correlation, but it's negative (-0.37). This means that a board game with a top bgg_rank (1) will most likely have more mechanics listed than a board game listed in the ranking list at 20,000, and the decrease in mechanic count will decrease as the rank decreases. So, yes, board game ranking, or popularity, does correlate with game mechanics listed.

The correlation test also confirms that this information is statistically significant due to a high t-value and a p-value that's nearly equivalent to 0.

Does a higher amount of mechanics also mean that the complexity rate will be higher? Are the more popular board games ones that are more complex or harder to complete?

```{r}

board_games_data_2 <- board_games_data |>
  mutate(mechanic_count = ifelse(
    is.na(mechanics) | mechanics == "",
    0,
    str_count(mechanics, ",") + 1
  ))

board_games_data_2 |>
  select(name, mechanics, mechanic_count, bgg_rank)

cor.test(board_games_data_2$bgg_rank, board_games_data_2$mechanic_count)
```

QUESTION: Which socioeconomic classes/education levels are more likely to have board games in their home? data set I found with details on this: <https://osf.io/nfr7d/files/tu8yq>

ANSWER: Individuals with higher educational attainment (university and postgraduate levels) are more likely to report engagement with board games and related hobby gaming behaviors. Lower educational groups (high school or some high school) show comparatively lower engagement levels.

```{r}

your_move_data <- read_sav("data/Your move data.sav") |> clean_names()
names(your_move_data)
your_move_data$education <- as_factor(your_move_data$education)
your_move_data$q3_1 <- as_factor(your_move_data$q3_1)
your_move_data$q4_8_6 <- as.numeric(your_move_data$q4_8_6)
```

```{r}
# 1. Education distribution

table(your_move_data$education)

ggplot(your_move_data, aes(x = education)) +
  geom_bar() + coord_flip() +
  labs(
    title = "Education Levels of Board Gamers",
    x = "Education Level",
    y = "Count"
  ) +
  theme_minimal()
```

```{r}
# 2. Education vs Player Type

ggplot(your_move_data,
       aes(x = education, fill = q3_1)) +
  geom_bar(position = "fill") + coord_flip() +
  labs(
    title = "Player Type by Education Level",
    x = "Education Level",
    y = "Proportion",
    fill = "Player Type"
  ) +
  theme_minimal()
```
