---
title: "Reverse Engineering Project"
author: "Adam and Logan"
date: "March 11"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, "Comic Books Are Still Made By Men, For Men And About Men"

<https://github.com/fivethirtyeight/data/tree/master/comic-characters>

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
dc_data <- read_csv("data/dc-wikia-data.csv") |> clean_names()
marvel_data <- read_csv("data/marvel-wikia-data.csv") |> clean_names()
```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: "Women and men were almost exactly equally likely to have a secret identity in the Marvel universe (49.4 percent of males and 49.5 percent of females), while in the DC universe 51 percent of males had a secret identity and 45 percent of females did.”
-   **Analysis summary**: Our analysis of DC characters did match the findings of the article at 45%. Our data did have an issue with 577 unclassified female characters, which could be relevant if the data is missing or not important if the characters do not have heroic/villainous identities. Marvel had a slightly different classification system for identities. We had a different result of 49.4% female secret identities when the five "known to authorities" characters were excluded but the results matched again when they were added in.

```{r}
# Put code to reverse engineer sentence here
dc_data |> filter(str_detect(sex, "Female")) |> group_by(id, sex) |> summarise(count=n())
# Display results of code below this codeblock
marvel_data |> filter(str_detect(sex, "Female")) |> group_by(id, sex) |> summarise(count=n())
```

### Sentence 2

-   **Sentence text**: "The same trend bears out when we look at characters with higher appearance counts. When we zero in on the 2,415 DC and 3,342 Marvel characters with gender data who appear at least 10 times — the set of characters that probably gives the most consistent view of what we see on the page — female characters make up only 30.9 percent of the DC universe and 30.6 percent of the Marvel universe."
-   **Analysis summary**: The article's conclusion of 30.9% of DC characters with 10 appearances being female is close but inaccurate to our findings, including scenarios where rounding would be used. We found 738 female characters, representing a close but ultimately higher 31.04%. For Marvel, we had a similar result with a roughly close number but a higher one again. We found 30.84% of Marvel characters with 10 appearances were female, beating out the article's smaller 30.6% result. The results being so close suggests that an extra character could have been in the article's calculations, potentially by doing a process of over 10 instead of greater than or equal to 10.

```{r}
dc_apps_over_ten <- dc_data |>
filter(appearances >= 10) |>
filter(sex == "Male Characters" | sex == "Female Characters") |>
group_by(sex) |>
summarise(count = n()) |>
mutate(percent = count / sum(count) * 100)

marvel_apps_over_ten <- marvel_data |>
filter(appearances >= 10) |>
filter(sex == "Male Characters" | sex == "Female Characters") |>
group_by(sex) |>
summarise(count = n()) |>
mutate(percent = count / sum(count) * 100)

# Display results of code below this codeblock
dc_apps_over_ten
marvel_apps_over_ten
```

### Sentence 3

-   **Sentence text**: "When we focus on only the most frequently appearing characters in each universe — the stars of the show — the data changes. Of characters with gender data and 100 or more appearances (294 DC characters and 414 Marvel characters), only 29.0 percent of DC’s are female, and a similar 31.1 percent of the Marvel crowd is."
-   **Analysis summary**: We found the same conclusion as the article for DC female characters with at least 100 appearances. Even with rounding considerations, the 29.01% in our findings would go down to 29%. Marvel did not have the same outcome for this calculation. We had a difference of almost 0.2% at 31.29% instead of the article's 31.1%. This difference could be from extra inclusions in the article's dataset with the appearance counter as the likeliest source for the decimal change. There are likely few characters with exactly 100 appearances, matching the little tweaks that could alter a result by such a small amount.

```{r}
dc_apps_over_100 <- dc_data |>
filter(appearances >= 100) |>
filter(sex == "Male Characters" | sex == "Female Characters") |>
group_by(sex) |>
summarise(count = n()) |>
mutate(percent = count / sum(count) * 100)

marvel_apps_over_100 <- marvel_data |>
filter(appearances >= 100) |>
filter(sex == "Male Characters" | sex == "Female Characters") |>
group_by(sex) |>
summarise(count = n()) |>
mutate(percent = count / sum(count) * 100)

# Display results of code below this codeblock
dc_apps_over_100
marvel_apps_over_100
```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
