---
title: "Reverse Engineering Project"
author: "Adam and Logan"
date: "March 11"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, "Comic Books Are Still Made By Men, For Men And About Men"

<https://github.com/fivethirtyeight/data/tree/master/comic-characters>

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
dc_data <- read_csv("data/dc-wikia-data.csv") |> clean_names()
marvel_data <- read_csv("data/marvel-wikia-data.csv") |> clean_names()
```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: "Women and men were almost exactly equally likely to have a secret identity in the Marvel universe (49.4 percent of males and 49.5 percent of females), while in the DC universe 51 percent of males had a secret identity and 45 percent of females did.”
-   **Analysis summary**: Our analysis of DC characters did match the findings of the article at 45%. Our data did have an issue with 577 unclassified female characters, which could be relevant if the data is missing or not important if the characters do not have heroic/villainous identities. Marvel had a slightly different classification system for identities. We had a different result of 49.4% female secret identities when the five "known to authorities" characters were excluded but the results matched again when they were added in.

```{r}
# Put code to reverse engineer sentence here
dc_data |> filter(str_detect(sex, "Female")) |> group_by(id, sex) |> summarise(count=n())
# Display results of code below this codeblock
marvel_data |> filter(str_detect(sex, "Female")) |> group_by(id, sex) |> summarise(count=n())
```

### Sentence 2

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
