---
title: "lab_11"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

- Our usual libraries for working with data, plus rvest.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Turn off scientific notation
options(scipen=999)

library(tidyverse)
library(janitor)
library(rvest)
library(lubridate)
```

Let's get to scraping.

## Questions

**Q1**. Scrape the listing of available Maryland physician disciplinary alerts at [www.mbp.state.md.us/disciplinary_2025.aspx](www.mbp.state.md.us/disciplinary_2025.aspx) into a dataframe. You should have three columns, one of which is a date, so make sure the date column has a date datatype. What's the most common sanction, and how many times is it mentioned in the 2024 alerts?

**A1** The most common sanction is Summary Suspension which appears nine times in the dataset. There were only six of those suspensions in 2024.

```{r}
physician_url <- "https://www.mbp.state.md.us/disciplinary_2025.aspx"
```

```{r}
physician <- physician_url |> read_html()
```

```{r}
phy_as_tbl <- physician |> html_table()
phy_as_tbl <- phy_as_tbl[[1]] |> mutate(Date=mdy(Date))
phy_as_tbl
```

```{r}
phy_as_tbl |> group_by(Sanction) |> summarise(count = n()) |> arrange(desc(count))
```

```{r}
alt_phys_url <- "https://www.mbp.state.md.us/disciplinary_2024.aspx"
```

```{r}
alt_physician <- alt_phys_url |> read_html()
```

```{r}
alt_phy_as_tbl <- alt_physician |> html_table()
alt_phy_as_tbl <- alt_phy_as_tbl[[1]] |> mutate(Date=mdy(Date))
alt_phy_as_tbl
```

```{r}
alt_phy_as_tbl |> filter(Sanction=="Summary Suspension.")
```

**Q2** Next, let's scrape the list of press releases from Maryland's Office of the Public Defender, <https://www.opd.state.md.us/press-releases>. This isn't a table, so you'll need to use `html_elements()` and your browser's inspector and do some clean up on the results. The result should be a dataframe with two columns that contain the date and title, and the date column should have a date datatype. The challenge here is figuring out how to isolate the releases.

When you finish scraping into a dataframe, write code to find the press releases that have the word "police" in the title.

**A2** There are nine press releases that have police in the title.

```{r}
opd_url <- "https://www.opd.state.md.us/press-releases"
```

```{r}
opd_read <- opd_url |> read_html()
```

```{r}
opd_read
```

```{r}
opd_as_tbl <- opd_read |> html_elements('p.wixui-rich-text__text') |> html_text() |> as_tibble() |> separate(value, c('date', 'title'), sep=":") |> drop_na(title) |> mutate(title = str_squish(title), date = mdy(date))
opd_as_tbl
```

```{r}
opd_as_tbl |> filter(str_detect(title, "Police")) |>nrow()
```

**Q3** Sen. Chris Van Hollen, D-Maryland, has posted press releases at <https://www.vanhollen.senate.gov/news/press-releases>. It would be great to have them in a dataframe that has the following columns: date, title and url.

To do this, you will need to scrape the page's html and save that to a variable, and *then* extract the dates, titles and urls into separate dataframes using html_elements(). There are two different elements that contain the press release URL; pick one of them, and you may need to invoke an HTML `class` value to do it. The function `html_text()` pulls out the contents of a tag, but for urls we want the HTML attribute. Rvest gives you a way to extract the URL from a link; google to find out what it is. Remember how we turn a list into a dataframe.

At the end, you'll have three dataframes that you want to combine into a single dataframe. When we want to combine the rows of identical dataframes, we used `bind_rows()`. If you were combining columns instead of rows, there's a similar function. Find out what it is and use it to put all of the dataframes together into a single one.

When you're done, rename the columns so they make sense, then make sure the date column is an actual date.

Finally, tell me what questions you could ask of this data if you had all of his press releases. Be creative.

**A3**: After looking at all of the urls, I am curious about the system for tagging representatives for articles. Van Hollen is not at the center of all of these, which suggests his staff are not assigning the articles and use an automated system. How can an article that is against Van Hollen get filtered out from his teams's website? If someone says his name in their quote, is that enough to get it to appear in the list? For the titles, what is the criteria for listing the names of the politicians instead of simplifying it as "colleagues" like some of the articles do? Does the algorithm for the website always append Van Hollen's name first before going alphabetically for the rest or is it another system? Why are the lengths of the titles so inconsistent and not standardized?

```{r}
van_hollen_url <- "https://www.vanhollen.senate.gov/news/press-releases"
```

```{r}
van_hollen <- van_hollen_url |> read_html()
```

```{r}
dates <- van_hollen |> html_elements("p.Heading") |> html_text()
dates
```

```{r}
urls <- van_hollen |> html_elements("a.ArticleBlock__moreLink") |> html_attr("href")
urls
```

```{r}
titles <- van_hollen |> html_elements("h2.ArticleTitle") |> html_text()
titles
```

```{r}
combined_van_hollen <- data.frame(Date = dates, Title = titles, Link = urls) |> mutate(Date = mdy(Date))
combined_van_hollen
```
