---
title: "lab_03"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TASK: Load libraries

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)
# Load janitor
library(janitor)
```
TASK: Load data and clean up the column names

Load the two datasets we worked with in the pre_lab that are in the data folder: the Maryland payments to vendors and Maryland grants and loans AND, using janitor, change the column names so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
md_st_payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv") |> clean_names()
# Load in the Maryland grants and loan data
md_grant_loan_payments <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv") |> clean_names()
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed in the payments data. Look at & search the records in the dataframe to start your search, starting with a broad search and then narrowing it down. Then write code to find rows where the vendor name contains some version of the county's school system.

#HINT: you should use filter() just once, and inside your filter you want to look for the variations one at a time. Think "this version OR that version."

#When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find?

A1: 98. I chose to include the board of education because they are also part of the school system. I chose school and education as other keywords because they are the most applicable and likeliest to be relevant for the search. These parameters could exclude other abbreviation methods or specific educational types that do not use either word.

```{r}
pg_schools <- md_st_payments |> filter(str_detect(vendor_name, "SCHOOL|EDUCATION") & str_detect(vendor_name, "PRINCE GEORGE|PG"))
nrow(pg_schools)
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? Are there any aspects of the results that stand out to you?

A2:2024 had the largest total amount. I find it interesting that there is no data from 2023 and that spending fell from the highest in 2024 to the lowest in 2025.

```{r}
pg_by_year <- pg_schools |> group_by(fiscal_year) |> summarise(total_amount = sum(amount)) |> arrange(desc(total_amount))
pg_by_year
```

Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received.

Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe in words how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3. I would clean the data by using a function to find similarities between different strings and renaming them when the percent that they are identical is over a high enough threshold. That would get rid of typo concerns or extra characters like unnecessary punctuation. The three examples of some vendor names that should be combined: "A MORTON THOMAS & ASSOC INC"	vs "A MORTON THOMAS & ASSOCIATES INC", "A&I INC" vs "A&I INC.", "AEROSOL MONITORING & ANALYSIS INC" vs "AERSOL MONITORING & ANALYSIS INC"

```{r}
umd_capital <-md_st_payments |> filter(str_detect(agency_name, "UNIVERSITY OF MARYLAND") & str_detect(agency_name, "CAPITAL PROJECT")) |> group_by(vendor_name) |> summarise(count = n()) #mutate(total_amount = sum(amount)) #|> group_by(vendor_zip) |> summarise(count_vendors = n())
umd_capital
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which column(s) contain the information you want, your code should find just the grants for stem cell research in a single filter. Then calculate the total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: Stem cell research funding has generally decreased year to year and especially dropped since 2013. COVID likely caused 2021 to have the lowest recorded funds. Spending skyrocketed from then on and the number reached the closest since 2009's leading value in 2023.

```{r}

stem_cell_research <- md_grant_loan_payments |> mutate(description = str_to_upper(description)) |> filter(str_detect(description, "STEM CELL")) |> group_by(fiscal_year) |> summarise(total_amount = sum(amount))
stem_cell_research
```

Q5: Using the Maryland grants and loan data, find records where the zip code is the one where the University of Maryland-College Park is located in and produce a list of grantees from that zip code, along with the total amount they received. Write a paragraph explaining how many grantees are there, what you would need to do to clean the results so that you had one record per grantee and which grantees are most interesting to you?

A5: There are 51 unique grantees listed in the dataset but that does not mean there are actually 51 different ones. Some grantees are only different because they have inconsistent spacing or punctuation. For example, both "COLLEGE PARK CITY - UNIVERSITY PARTNERSHIP INC" and "COLLEGE PARK CITY-UNIVERSITY PARTNERSHIP" are separated as different grantees in the dataset. To clean the data further, the strings would all need to be normalized through uppercasing or lowercasing and then a string similarity function could identify nearly identical strings. The most interesting ones included are the companies like Ikea and HAIR & SPACE BLOWDRY & BEAUTY CO which do not match the university or city owned entries.

```{r}
umd_zips <- md_grant_loan_payments |> filter(str_detect(zip_code, "20740|20741|20742")) |> group_by(grantee) |> summarise(total_amount = sum(amount)) #|> nrow()
umd_zips
```
