---
title: "lab_12"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

- tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
library(rvest)
```

## Questions

**Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

**A1.** The leading word after getting rid of the stop words was "failing." The top 10 words all seem to be describing issues between clients paying their attorneys and dissatisfaction from lost cases.

```{r}
sanctions <- read_csv("data/md_attorney_sanctions.csv")
```

```{r}
sanction_unique <- sanctions |> select(text) |> unnest_tokens(word, text)
sanction_unique
```

```{r}
sanction_unique |>
  count(word, sort = TRUE) |>
  top_n(25) |>
  mutate(word = reorder(word, n)) |>
  ggplot(aes(x = word, y = n)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
      labs(x = "Count",
      y = "Unique words",
      title = "Count of unique words found")
```

```{r}
data("stop_words")

stop_words <- stop_words |>
  add_row(word = "to") |> 
  add_row(word = "the") |> 
  add_row(word = "for") |>
  add_row(word = "and") |>
  add_row(word = "of") |>
  add_row(word = 'in') |> 
  add_row(word = 'his') |> 
  add_row(word = 'a') |>
  add_row(word = 'on') |>
  add_row(word = 'with') |>
  add_row(word= 'by') |>
  add_row(word = 'that') |>
  add_row(word = 'an') |>
  add_row(word = 'or')
```

```{r}
sanction_unique |>
  anti_join(stop_words) |>
  group_by(word) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(50)
```

**Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

**A2.** "For failing to" is the most common trigram and it appears 928 times in the sanctions data. It likely refers to a breach of contract between the client and the lawyer since the client would only file a sanction if their lawyer did their job improperly. If the stop words were removed, the highest trigram frequency would be "attorney trust account" and that refers to the management of client funds, which was wrongly done in this case.

```{r}
trigrams <- sanctions |> 
  select(text) |> 
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |> 
  group_by(trigram) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(10)
```

**Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

**A3.** The code suggests that negligence was not a significant factor in sanction worthy behavior. Most of the cases instead had criminal actions behind them. Negligence was largely consistent between 44 and 48 cases per year except for outliers in 2011 and 2012 at the start of the dataset. Spikes in 2015 and 2017 became outliers since they dropped in the following years. One explanation for that could be that lawyers who got probation sanctions finished their punishment and returned to the same behaviors. To clarify the reasoning for the sanctions, the text could be filtered to have a four word phrase that directly follows "for" or "to" since the sanctions generally follow that structure of the reasoning coming right after those words.

```{r}
sanctions |> filter(str_detect(text, "for failing to")) |> group_by(fiscal_year) |> summarise(count = n())
```
