---
title: "week7_recap"
author: "Daniel Trielli"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load Data

Run the codeblock below to load the data.

```{r}
earthquakes <- read_csv('https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv')

#Setting time column as datetime
earthquakes <- earthquakes |> mutate(time = as_datetime(time))
```

#### Answer the questions below

Most questions have a code block and a space for an answer below. Write the code you think is necessary and, in the answer space, write out what you did and what was the result.

------------------------------------------------------------------------

#### **Q1** Do you notice anything different with this read_csv()? Particularly in the file name?

**A1: The read_csv uses a url for the file instead of having it downloaded as a separate file.**

------------------------------------------------------------------------

#### **Q2** How many records there are there in this dataset? What do they mean and what useful information we can gather from it, looking at the columns?

**A2: There are 10,307 records but the dataset updates every few minutes with new entries.**

------------------------------------------------------------------------

#### **Q3** How do I reorganize this list to see the ones that are the deepest first? What is the depth that shows up for the deepest one, and its magnitude?

```{r}
deep_earthquakes <- earthquakes |> arrange(desc(depth))
deep_earthquakes
```

**A3: The depth of the deepest earthquake is 667.197 and it has a 4.30 magnitude.**

------------------------------------------------------------------------

#### **Q4** I just want to see the earthquakes with a magnitude larger than 6. How do I do that? And how many are there that fit this criteria?

```{r}
over_six <- earthquakes |> filter(mag > 6)
nrow(over_six)
```

**A4: There are nine earthquakes with a magnitude over six.**

------------------------------------------------------------------------

#### **Q5** What about if I want to see earthquakes that have both a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [both]{.underline} these criteria?

```{r}
over_six_and_below_20 <- over_six |> filter(depth < 20)
nrow(over_six_and_below_20)
```

**A5: There are four earthquakes that have a magnitude over 6 and a depth under 20.**

------------------------------------------------------------------------

#### **Q6** What about if I want to see earthquakes that either have a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [either]{.underline} these criteria?

```{r}
over_six_or_below_20 <- earthquakes |> filter(depth < 20 | mag > 6)
nrow(over_six_or_below_20)
```

**A6: There are 8,250 earthquakes that either have a depth under 20 or a magnitude over 6.**

------------------------------------------------------------------------

#### **Q7** I'm interested in finding earthquakes that took place in Alaska. Which column should I look at? How do I use it to find all the earthquakes in Alaska? How many earthquakes took place there?

```{r}
in_alaska <- earthquakes |> filter(str_detect(place, 'Alaska'))
nrow(in_alaska)
```

**A7: There are 3,327 earthquakes that happened in Alaska.**

------------------------------------------------------------------------

#### **Q8** I notice that there is a column called 'type', that seems to have different kinds of tremors. What is the best way to find what are all the possible types of tremors, and counting how many of each there are in this data set? What are the first two most common types of tremors in this data set?

```{r}
earthquake_types <- earthquakes |> group_by(type) |> summarise(count_types = n()) |> arrange(desc(count_types))
earthquake_types
```

**A8: The two most common earthquake types are earthquakes and quarry blasts.**

------------------------------------------------------------------------

#### **Q9** How do I calculate the average depth of each type of tremor? What is the average depth of the earthquake type in this data set? Is there anything that seems unusual or surprising in this finding?

```{r}
earthquake_avg_depth <- earthquakes |> group_by(type) |> summarise(avg_depth = mean(depth))
earthquake_avg_depth
```

**A9: The average depth of the earthquake type is 21.86. The data sticks out for being significantly larger than any other type, going over the second highest in mining explosion by 15.**

------------------------------------------------------------------------

#### **Q10** I'm interested, in the future, to see the hours in which earthquakes happen. How can I extract the hour from the time column?

```{r}
e_by_hour <- earthquakes |> mutate(hour = hour(time))
e_by_hour
```

**A10: Hours can be extracted from a date with the hour() function which is applied to the time column.**

------------------------------------------------------------------------

#### **Q11** I want to make a note of all the records in this data set that I consider serious. For my purposes, I'm thinking that all tremors that have a magnitude that is larger than 3 are serious. How do I automatically create a new column making that distinction?

```{r}
with_serious <- e_by_hour |> mutate(serious = case_when(
  mag > 3 ~ 'Yes',
  .default = 'No'
))
with_serious
```

**A11: The case_when() function allows for values to be assigned in a new column based on data from another column.**

------------------------------------------------------------------------

#### **Q12** I have no idea how earthquakes work and I'm interested in seeing if there is a particular time of day in which serious earthquakes happen. How can I see that condensed in a table with all the hours in a day and all the serious earthquakes in each hour? What is the hour with fewer serious earthquakes and the one with the most serious earthquakes?

```{r}
serious_by_hour <- with_serious |> filter(str_detect(serious, "Yes")) |> group_by(hour) |> summarise(count_serious = n()) |> arrange(desc(count_serious))
serious_by_hour
```

**A12**: 5:00 p.m. (17) is the hour with the most serious earthquakes. Noon (12) is the hour with the fewest serious earthquakes.

------------------------------------------------------------------------

#### EXTRA Q What other questions are you interested in and how would you ask them in R?

My first question is how many serious earthquakes are in the morning vs after noon.

My second question is about the most common direction from a listed place in the data.

My third question is how earthquake type related to the depth error.

```{r}
serious_by_hour |> mutate(time_of_day = case_when(
  hour > 12 ~ 'After Noon',
  .default = 'Before Noon')) |> group_by(time_of_day) |> summarise(count = n()) |> arrange(desc(count))
```

```{r}
with_serious |> mutate(direction = case_when(
  str_detect(place, " E ") ~ 'East',
  str_detect(place, " N ") ~ 'North',
  str_detect(place, " S ") ~ 'South',
  str_detect(place, " W ") ~ 'West',
  str_detect(place, "NW") ~ 'Northwest',
  str_detect(place, "NE") ~ 'Northeast',
  str_detect(place, "SW") ~ 'Southwest',
  str_detect(place, "SE") ~ 'Southeast',
  .default ='unknown')) |> group_by(direction) |> summarise(count = n()) |> arrange(desc(count))
```

```{r}
earthquakes |> group_by(type) |> summarise(avg_depth_error = mean(depthError))
```
